---
title: "(avocado)"
author: "Luidmila Filippova"
date: "11/21/2024"
format:
  html:
    embed-resources: true
editor: source
---

````markdown
```{r setup, message=FALSE, warning=FALSE, include=FALSE}
library(dplyr) 
library(readr) 
avocado_data <- read_csv("avocado.csv")
```

### Question 1: What is the average price difference between organic and conventional avocados?

```{r}
avocado_data %>% 
  group_by(type) %>% 
  summarise(Avg_Price = mean(AveragePrice, na.rm = TRUE)) %>% 
  arrange(Avg_Price)
```

### Question 2: Which regions had the highest total volume of sales in 2018?

```{r}
avocado_data %>% 
  filter(year == 2018) %>% 
  group_by(region) %>% 
  summarise(Total_Volume = sum(`Total Volume`, na.rm = TRUE)) %>% 
  arrange(desc(Total_Volume)) %>% 
  head(5)
```

### Question 3: Calculate the percentage of Total Volume represented by Small Bags.

```{r}
avocado_data %>% 
  mutate(Small_Bag_Percent = (`Small Bags` / `Total Volume`) * 100) %>% 
  select(region, Date, Small_Bag_Percent) %>% 
  arrange(desc(Small_Bag_Percent))
```

### Question 4: List the Average Price and Total Volume for all organic sales in New York.

```{r}
avocado_data %>% 
  filter(region == "NewYork", type == "organic") %>% 
  select(Date, AveragePrice, `Total Volume`) %>% 
  arrange(desc(Date))
```

### Question 5: Find the region with the lowest average price across all years.

```{r}
avocado_data %>% 
  group_by(region) %>% 
  summarise(Overall_Avg_Price = mean(AveragePrice, na.rm = TRUE)) %>% 
  arrange(Overall_Avg_Price) %>% 
  head(1)
```

```
```